---
title: "excise report"
author: "Krisna Gupta"
institute: "Prospera"
date: last-modified
date-format: long
bibliography: ref.bib
abstract: "This report presents our analysis on mainly estimating elasticity and forecasting future excise and revenue for DGCE. We find that the elasticity of demand to be vary between -0.5 to -1.1. We also find a strong case for tax-price pass-through for SKM and SPM tobacco type. Our findings corroborates most literature on the Indonesian tobacco use. Unfortunately data limits how much and how robust the analysis can get. In particular, we find very little value of analysis on electric cigarettes, which maindly due to only one year datapoint available for study."
abstract-title: "executive summary"
format:
  html:
    code-fold: true
  pdf:
    documentclass: scrreprt
    papersize: a4
  docx:
    toc: false
---

## Introduction {#sec-introduction}

Nicotine is addictive and harmful. Controlling tobacco use via taxation requires understanding of demand elasticity of the good [@RN]. ADBI was tasked to help Min o Finance to calculate the elasticity of cigarette excise.

The study become even more important since the introduction of electric cigarettes. It can be argued that e-cigs create a substitute to the cigarettes, and may be an introduction to more traditional tobacco cigarettes [@binns]. DGCE was starting to collect data (and excise) for e-cigarettes since 2020.

We must answer this following question: How much is the cross-price substitution between traditional cigarettes?

While all of these questions are extremely important, we find the answer to not be straight-forward. Few literatures actually looking at this. But the main problem is the lack of good quality data in answering those questions, particularly on e-cigarettes.

## Literatures

We rely on three papers for our desk investigation which is summarized in @tbl-1.

@RN investigates whether smokers in general are myopic or rational addicts. They utilise 2SLS and GMM along with IFLS data with 1783 total observations to conduct their study. According to this consumer-side study, the price elasticity of demand for traditional tobacco is around -0.38 to -0.57. They also controls for income with total expenditure as the proxy.

@nach look at the impact of cigarette prices to household nutrient intake. Collecting various household surveys in 2014, they are able to observe a cross-section 285,400 households. They use unit value as the price which is sourced from PODES, one of the household survey. While their goal is not to look at elasticity, they find that 1% increase in cigarette prices lead to an increase of budget share for cigarettes by 0.0737 percentage point.

@prasetyo618 utilise transactional data on the firm level, 2005-2017, with total 32,711 observations (around 2,500 firms per year). They get the data from DGCE. They look at the tax-price pass through, which is the impact of changes in tax to the firms' pricing strategy. They also examine the pass-through effect on three types of traditional cigarettes, namely SKT, SKM and SPM. They find that an increase of 1% of tax rate leads to an increase of prices by around 0.15, 0.36 and 0.77 for SKT, SKM and SPM respectively.

| paper  | finding  | data | obs year | N |
|--------|--------|------|------|-----|
| @RN  |  elasticity (-0.38 to -0.57)  | IFLS | 1993,1997,2000 | 1783 |
| @nach   | budget share   | Susenas, PODES, RISKESDAS  | 2014 | 285,400 |
| @prasetyo618 | tax-price passtrough   | firm transactions from DGCE | 2005-2017 | 32,711 |


: literature summary {#tbl-1 .striped .hover}

For objectives, this report bear similarity with @RN the most, albeit it is indeed important to see a tax-price pass-through a la @prasetyo618. We do have a production data of different types of cigarettes, but our data lack firm level richness as @prasetyo618[^2]. Generally we do not look at the question posed by @nach, but This is still an important question to adhere by DGCE. The reason is because using excise as an instrument to reduce smoking may lead to a reduction of a household's nutrient intake.

@prasetyo618 offers various insight to the cigarettes market in Indonesia. The industry is highly oligopolistic and often offers low price of cigarettes to avoid paying progressive exicse. This leads to higher number of smokers[^3]. Low potential revenue,thus may be outweighed by the higher cost of treatment of tobacco-related diseases.

On the last note, none of those are looking at electric cigarettes.

[^2]: interestingly, they also sourced the data from DGCE.
[^3]: adult male smokers increase in share from 60.6% in 2000 to 76.1%

## Data

This paper relies solely on data sourced by DGCE. The data contains information on both traditional and e-cigarettes, albeit on different details. We will first discuss about the traditional cigarettes and then proceed to e-cigarettes.

### Traditional cigarette data

The traditional cigarettes dataset contains information on annual production (in unit) and excise revenue for three kinds of traditional cigarettes, SKM, SKT and SPM. It also contain two types of regulated price data, HTP (retail price) and HJE (base price to calculate the excise liabilities), both quarterly. The data is available from 2018-2022, 2023 data was available for the first and second quarter only.

```{r}
#| echo: false
#| warning: false
#| message: false
#| error: false
library(tidyverse)
library(readxl)
library(modelsummary)
library(kableExtra)
library(gt)
library(RStata)

theme_set(theme_classic())


```

```{r}
options("RStata.StataPath"="\"C:\\Program Files\\Stata17\\StataMP-64\"")
options("RStata.StataVersion" = 17)

## Reading data and write some logs
dat<-stata('import excel "data\\excisedata.xlsx", sheet("Sheet1") firstrow',data.out=TRUE)

stata('sureg (lqskm1 llskm1 llskm2 llspm1 llspm2 llskt1a llskt1b llskt2 llskt3) (lqskm2 llskm1 llskm2 llspm1 llspm2 llskt1a llskt1b llskt2 llskt3)',data.in = dat)

stata('twoway (line qskm1 mo) (line qskm2 mo)',data.in = dat)
```

WHO suggests any tax rate below 70% is deemed too small to reduce smoking [@prasetyo618]



## Method

Here we describe our method in getting elasticities, tax-price pass-through and marginal revenue per marginal tax increase, given the data limitation.

Let a standard downward-sloping demand function:

$$
Q_i=AP_i^{-\varepsilon}
$$

which can be approximately log-linearized

$$
q_i=a-\varepsilon p_i
$$

where a lowercase is the log version of its uppercase counterparts. We can, thus, econometrically estimated the above equation with a regression. Additionally, we follow the theory in having an income elasticity with GDP per capita as a proxy. We then estimate:

$$
q_{i}=a-\varepsilon p_{i} + \gamma y_{i} \pm \epsilon_{i}
$$

We assume an iid $\epsilon_{it}$ for now and uses own-price elasticity since we lack information on the price of electric cigs. The parameter $\varepsilon$ is the own-price elasticity of demand, which we expect to be negative, while $\gamma$ is the income elasticity of demand which is assumed to be positive.

We have index $i \in {\text{traditional},\text{SKM},\text{SKT},\text{SPM}}$. That is, we will regress on the level of each type of traditional cigarettes. But we will also regress on aggregate. That is, we regress total demand of traditional cigarettes $Q_{traditional}=\sum_i Q_i$ against a weighted average prices:

$$
P_{traditional}=\frac{\sum_i P_i Qi}{\sum_i Q_i} \ \ for \ i \in \{SKM,SPM,SKT\}
$$

## Results

We use data that are made available by DGCE for us. Data contains production, revenue, base price (HJE) for tax purposes and retail price (HTE) for consumers for various types of traditional and electric cigarettes. Unfortunately, the availability of those data differs quite tremendously. Some are collected monthly, some others are quarterly and annually. Common denominator suggests us to conduct the analysis annually.



## References {.unnumbered}


